---
title: "TreesMeta_Old_New_Separate"
author: "Ryan Fass"
date: "2023-01-18"
output: html_document
---
### Laguna Mountain: Tables of Alive, Dead and Topkill Adult QUKEs and Bar Graphs. Old Experiment and New Experiment separate. 



TreesMeta_Old_New_Separate means: same as tree data, where I separated the old and new experiments into two different csv’s so that processing is a little more streamline. I separated the TreesMeta_data in Excel. For Old Experiment: I deleted all of the New Experiment plots and the two extra controls (190 & 191), using Sort in excel. For New Experiment: I only added Old Experiment Control plots (PostBurnYear1 and changed to PreBurn) and (PostBurnYear2 and changed to PostBurnYear1) This change in monitoring_status was to mimic the year of survey for the New Experiment.

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(here)
library(janitor)
library(knitr)
library(dplyr)
library(patchwork)
```


Read in the two tree csv's (old and new) 
```{r}
trees_all_old_data <- read_csv(here('trees_all_old_data.csv'))
trees_all_new_data <- read_csv(here('trees_all_new_data.csv'))
```


Take both the trees old experiment data and the tree new experiment data and select just the QUKE species.Then change the status_x values (L = live and D = dead). Create a new column call final_status that has values = alive, dead, topkill. 
```{r}
trees_tk_status_old <- trees_all_old_data %>%
  select(macro_plot_name, monitoring_status, status_x, species_symbol,  treatment_burn,  experiment, top_kill_y_n) %>%
  filter(species_symbol %in% c("QUKE")) %>%
  mutate(live = status_x == "L") %>%
  mutate(dead = status_x == "D",
         final_status = case_when(live == TRUE ~ "alive",
                                  dead == TRUE & top_kill_y_n == "N" ~ "dead",
                                  dead == TRUE & top_kill_y_n == "Y" ~ "topkill",
                                  dead == TRUE & is.na(top_kill_y_n) == TRUE ~ "dead"))

trees_tk_status_new <- trees_all_new_data %>%
  select(macro_plot_name, monitoring_status, status_x, species_symbol,  treatment_burn,  experiment, top_kill_y_n) %>%
  filter(species_symbol %in% c("QUKE")) %>%
  mutate(live = status_x == "L") %>%
  mutate(dead = status_x == "D",
         final_status = case_when(live == TRUE ~ "alive",
                                  dead == TRUE & top_kill_y_n == "N" ~ "dead",
                                  dead == TRUE & top_kill_y_n == "Y" ~ "topkill",
                                  dead == TRUE & is.na(top_kill_y_n) == TRUE ~ "dead"))
```



For the Old and New experiments, I created a data frame that has the columns that I want to show in the table (using group_by) and then used the summarise fuction to count the Alive, Dead and Topkill in the final status.
```{r}
group_tk_old = trees_tk_status_old %>% 
  group_by(monitoring_status, treatment_burn, experiment) %>%  
  summarise(live = sum(final_status== "alive"),
            dead = sum(final_status== "dead"),
            topkill = sum(final_status== "topkill")) %>% 
 mutate(monitoring_status = as.factor(monitoring_status))
```

Reordering the data frame so that the monitoring_status reads logically. Then using ‘kable’ to create the tables.
```{r}
group_tk_old_reorder <- group_tk_old %>% 
  mutate(monitoring_status=fct_relevel(monitoring_status, "PreBurn", "PostBurnYear1", "PostBurnYear2")) %>%
  arrange(monitoring_status)# Arranges the data frame by this variable
  
kable(group_tk_old_reorder, format = "markdown", col.names = c("Monitoring Status", "Treatment", "Experiment", "Alive", "Dead", "Topkill"), align = "llcccc", digits = 3,caption = "Number of Alive, Dead and Topkill Adult QUKEs by burn treatment")


write.csv(group_tk_old_reorder, here("group_tk_old_reorder.csv"))
```

```{r}
group_tk_new = trees_tk_status_new %>% 
  group_by(monitoring_status, treatment_burn, experiment) %>%  
  summarise(live = sum(final_status== "alive"),
            dead = sum(final_status== "dead"),
            topkill = sum(final_status== "topkill")) %>% 
 mutate(monitoring_status = as.factor(monitoring_status))
```

```{r, warning=FALSE}
group_tk_new_reorder <- group_tk_new %>% 
  mutate(monitoring_status=fct_relevel(monitoring_status, "PreBurn", "PostBurnYear1", "PostBurnYear2")) %>%
  arrange(monitoring_status)# Arranges the data frame by this variable
  
kable(group_tk_new_reorder, format = "markdown", col.names = c("Monitoring Status", "Treatment", "Experiment", "Alive", "Dead", "Topkill"), align = "llcccc", digits = 3,caption = "Number of Alive, Dead and Topkill Adult QUKEs by burn treatment") 

```

```{r}

plot_tk_status_new <- group_tk_new_reorder %>%
  pivot_longer(cols = c('live', 'dead', 'topkill'), 
               names_to = 'status',
               values_to = 'count')


ggplot(data = plot_tk_status_new, aes(fill = status, x = treatment_burn, y = count))+
  geom_bar(stat= "identity", 
           position= "stack",
           color="black")+
  scale_fill_manual(values = c('chocolate', 'darkgreen', 'green'))+
 # geom_text(aes(y = 42,label= treatment_burn),
#            position = position_dodge(width = .9),
 #           angle = 90,
 #           fontface= "bold")+
  facet_wrap(~monitoring_status)+
  ylab('Number of Trees') +
   xlab('Treatment')+
  ggtitle("New Experiment- Adult Quercus kelloggii ")+
  theme_classic()+
  theme(axis.text.x = element_text(size =9, angle = 25, hjust =1))
```

```{r}
plot_tk_status_old <- group_tk_old_reorder %>%
  pivot_longer(cols = c('live', 'dead', 'topkill'), 
               names_to = 'status',
               values_to = 'count')


ggplot(data = plot_tk_status_old, aes(fill = status, x = treatment_burn, y = count))+
  geom_bar(stat= "identity", 
           position= "stack",
           color="black")+
   scale_fill_manual(values = c('chocolate', 'darkgreen', 'green'))+
 # geom_text(aes(y = 30,label= treatment_burn),
 #           position = position_dodge(width = .9),
 #           angle = 90,
  #          fontface= "bold")+
  facet_wrap(~monitoring_status)+
  ylab('Number of Trees') + 
  xlab('Treatment')+
  ggtitle("Adult Quercus kelloggii- Old Exp.")+
  theme_classic()+
  theme(axis.text.x = element_text(size =9, angle = 25, hjust =1))

ggsave(here::here("figures", "oldexp_adults_count.jpg"),
       device = "jpg")
```


GGPlot of Old Experiment with Dead before burn taken out!
```{r}
#Reading in new csv with intial dead subtracted out in Excel
oldexp_adult_no_dead <- read_csv(here('group_tk_old_no_dead.csv'))


#ggplot

oldexp_adult_no_dead_reorder <- oldexp_adult_no_dead %>% 
  mutate(monitoring_status=fct_relevel(monitoring_status, "PreBurn", "PostBurnYear1", "PostBurnYear2")) %>%
  arrange(monitoring_status)# Arranges the data frame by this variable

plot_status_no_dead <- oldexp_adult_no_dead_reorder %>%
  pivot_longer(cols = c('live', 'dead', 'topkill'), 
               names_to = 'status',
               values_to = 'count')


ggplot(data = plot_status_no_dead, aes(fill = status, x = treatment_burn, y = count))+
  geom_bar(stat= "identity", 
           position= "stack",
           color="black")+
   scale_fill_manual(values = c('chocolate', 'darkgreen', 'green'))+
 # geom_text(aes(y = 30,label= treatment_burn),
 #           position = position_dodge(width = .9),
 #           angle = 90,
  #          fontface= "bold")+
  facet_wrap(~monitoring_status)+
  ylab('Number of Trees') + 
  xlab('Treatment')+
  ggtitle("Adult Quercus kelloggii- Old Exp.")+
  theme_classic()+
  theme(axis.text.x = element_text(size =9, angle = 25, hjust =1))

ggsave(here::here("figures", "oldexp_adults_count_no_dead.jpg"),
       device = "jpg")
```







